HEP Software Foundation Community White Paper Working Group -- Detector Simulation

### 2018-03-01 CWP/papers/HSF-CWP-2017-07_detsim/latex/DetectorSimulation.tex
final corrections to references and subsection numbering

Components of the HEP simulation chain, such as physics generators and
event visualisation tools are discussed in other chapters, while the
modeling of the propagation and beam manipulation through accelerator
lattices (i.e. CERN and Fermilab accelerator complexes) are considered
to be outside the scope of this chapter. Space and medical applications
are referred to briefly where appropriate.

@@

Electromagnetic (EM) transport simulation is challenging as it occupies
a significant part of the computing resources used in full detector
simulation. More experimental data in the coming decade implies higher
statistics and this implies the need for large and accurate simulation
samples, which require both model accuracy and speed. Therefore the
physics models need to be improved for achieving higher accuracy and the
implementation of algorithms needs to be reviewed in order to optimise
performance. Significant efforts have been made in the recent past to
better describe the simulation of electromagnetic shower shapes, in
particular to model the H $\rightarrow \gamma \gamma $ 
signal accurately at the LHC. This effort is being continued with 
emphasis on reviewing the
models assumptions, approximations and limitations, especially at very
high energy, and with a view to improving their respective software
implementations. For LHC Run-1 simulation Geant4 version 9.4 was used
~\cite{1742-6596-331-3-032029}. For Run-2 significant improvements were introduced into the
bremsstrahlung, pair production and multiple scattering 
~\cite{1742-6596-396-2-022013} ~\cite{1742-6596-513-2-022015}
further improving simulation of the EM shower shape; these improvements
were made available in Geant4 9.5 and 9.6. For validation purposes, an
extended set of tests versus data has been developed for high, moderate
and low-energies using test-beam data and published thin target data.
Comparisons with low-energy EM models from the Livermore and Penelope
packages were also carried out.

The process of multiple scattering of charged particles is a key
component of Monte Carlo transport codes. At high energy, it defines the
deviation of charged particles from ideal tracks, limiting the spatial
resolution of detectors. The scattering of low energy electrons defines
the energy flow across volume boundaries. This affects the sharing of
energy between absorbers and sensitive elements, directly affecting
shower shapes. A comprehensive description of recent improvements of the
Geant4 electromagnetic module can be found in ~\cite{1742-6596-513-2-022015}. Rather good
agreement was found when Geant4 predictions were compared with
experimental data collected at the LHC.

@@

8.2 ATLAS and CMS

@@

\begin{itemize}
\item
  Machine: olhswep16.cern.ch (CERN's OpenLab), one thread runs
\end{itemize}

\begin{itemize}
\item
  Compiler: gcc 6.3
\item
  Geant4: version 10.2, FTFP\_BERT physics list
\item
  Pythia ~\cite{Sjostrand:2014zea} events: 13 TeV Pythia minimum bias (300 events) and
  ttbar (300 events), pseudo rapidity cut \textbar{}$\eta$\textbar{}\textless{}5.5
\item
  Particle gun: 50 GeV e's, muons, pions with a flat $\eta,\varphi$ distribution in
  $\eta={[}-0.8,0.8{]}, \eta={[}2,2.7{]}, and \varphi= {[}0,2\pi{]}$
\item
  Geometry: 2015 or 2016
\end{itemize}

>>>>

Components of the HEP simulation chain, such as physics generators and
event visualisation tools are discussed elsewhere, while the
modeling of the propagation and beam manipulation through accelerator
lattices (i.e. CERN and Fermilab accelerator complexes) are considered
to be outside the scope of this report. Space and medical applications
are referred to briefly where appropriate.

@@

Electromagnetic (EM) transport simulation is challenging as it occupies
a significant part of the computing resources used in full detector
simulation. More experimental data in the coming decade implies higher
statistics and this implies the need for large and accurate simulation
samples, which require both model accuracy and speed. Therefore the
physics models need to be improved for achieving higher accuracy and the
implementation of algorithms needs to be reviewed in order to optimise
performance. Significant efforts have been made in the recent past to
better describe the simulation of electromagnetic shower shapes, in
particular to model the H $\rightarrow \gamma \gamma $ 
signal accurately at the LHC. This effort is being continued with 
emphasis on reviewing the
models assumptions, approximations and limitations, especially at very
high energy, and with a view to improving their respective software
implementations. For LHC Run-1 simulation Geant4 version 9.4 was used
~\cite{1742-6596-331-3-032029}. For Run-2 significant improvements were introduced into the
bremsstrahlung, pair production and multiple scattering
further improving simulation of the EM shower shape
~\cite{1742-6596-396-2-022013}~\cite{1742-6596-513-2-022015}; these improvements
were made available in Geant4 9.5 and 9.6. For validation purposes, an
extended set of tests versus data has been developed for high, moderate
and low-energies using test-beam data and published thin target data.
Comparisons with low-energy EM models from the Livermore and Penelope
packages were also carried out.

The process of multiple scattering of charged particles is a key
component of Monte Carlo transport codes. At high energy, it defines the
deviation of charged particles from ideal tracks, limiting the spatial
resolution of detectors. The scattering of low energy electrons defines
the energy flow across volume boundaries. This affects the sharing of
energy between absorbers and sensitive elements, directly affecting
shower shapes. A comprehensive description of recent improvements of the
Geant4 electromagnetic module can be found in~\cite{1742-6596-513-2-022015}. Rather good
agreement was found when Geant4 predictions were compared with
experimental data collected at the LHC.

@@

\hypertarget{ATLAS-and-CMS}{%
\subsection{ATLAS and CMS}\label{ATLAS-and-CMS}}

@@

\begin{itemize}
\item
  Machine: olhswep16.cern.ch (CERN's OpenLab), one thread runs
\item
  Compiler: gcc 6.3
\item
  Geant4: version 10.2, FTFP\_BERT physics list
\item
  Pythia ~\cite{Sjostrand:2014zea} events: 13 TeV Pythia minimum bias (300 events) and
  ttbar (300 events), pseudo rapidity cut \textbar{}$\eta$\textbar{}\textless{}5.5
\item
  Particle gun: 50 GeV e's, muons, pions with a flat $\eta,\varphi$ distribution in
  $\eta={[}-0.8,0.8{]}, \eta={[}2,2.7{]}, and \varphi= {[}0,2\pi{]}$
\item
  Geometry: 2015 or 2016
\end{itemize}
